---
title: "LAGOS Spatial Analysis"
author: "Matthew Ross"
date: "9/11/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)
```{r data-read}
# #Lagos download script
# lagosne_get(version='1.087.1',dest_folder = LAGOSNE:::lagos_path())

#run this before Th
lagosne_get(dest_folder = LAGOSNE:::lagos_path())

  
# #Load in lagos
# lagos <- lagosne_load()
# 
# #Grab the lake centroid info
# lake_centers <- lagos$locus

load('lake_centers.Rdata')
```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100)

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r, echo=F }
Iowa_Illinois <- states %>%
  filter(name %in% c('Iowa','Illinois')) %>%
  st_transform(2163)


mapview(Iowa_Illinois)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa
combined? How does this compare to Minnesota?

### Illinois and Iowa have 16,466 sites. Minnesota has 29,038 sites.

```{r, echo=F}

#Subset lakes based on spatial position
ii_lakes <- spatial_lakes[Iowa_Illinois,]

str(ii_lakes)
str(minnesota_lakes)

```


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r, echo=F}

ii_lakes_name <- ii_lakes%>%
  mutate(state='Iowa_Illinois')

mn_lakes_name <- minnesota_lakes%>%
  mutate(state='Minnesota')

long_lake <- rbind(ii_lakes_name, mn_lakes_name)

#ggplot(ii_lakes, aes(lake_area_ha))+
  #geom_histogram(bins= 30)+
  #scale_x_log10()
  
#ggplot(minnesota_lakes, aes(lake_area_ha))+
  #geom_histogram(bins= 30)+
  #scale_x_log10()

ggplot(long_lake, aes(lake_area_ha))+
  geom_histogram(bins= 30)+
  scale_x_log10()+
  labs(x='Lake Area (Ha)', y='Frequency')+
  facet_wrap(~state, dir="v")

```


## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r, echo=F}

#Plotting the first 1000 lakes
ii_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')

```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 

### Other sources of data that we might use to explore the difference in lakes 
and reservoirs in the three starts are 
