---
title: "NE Lakes and Spatial Intro"
author: "Danielle Reimanis"
date: "9/18/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
```

```{r load lagos, echo=F, warning=F}
#LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())

load('lake_centers.Rdata')
```


```{r, echo=F, results=F}
#turn to spatial data
#don't have lake_centers
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)

states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')


```

# In-Class work


### 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r, echo=F }
Iowa_Illinois <- states %>%
  filter(name %in% c('Iowa','Illinois')) %>%
  st_transform(2163)


mapview(Iowa_Illinois)

```



### 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

Illinois and Iowa have 16,466 sites. Minnesota has 29,038 sites.

```{r, echo=F}

#Subset lakes based on spatial position
ii_lakes <- spatial_lakes[Iowa_Illinois,]

str(ii_lakes)
str(minnesota_lakes)

```


### 3) What is the distribution of lake size in Iowa vs. Minnesota?

```{r, echo=F}

iowa_mn <- states %>%
  filter(name %in% c('Iowa','Minnesota')) %>%
  st_transform(2163)

iowa_mn_lakes <- spatial_lakes %>%
  st_join(.,iowa_mn)%>%
  filter(!is.na(name))

ggplot(iowa_mn_lakes, aes(lake_area_ha))+
  geom_histogram(bins= 30)+
  scale_x_log10()+
  labs(x='Lake Area (Ha)', y='Frequency')+
  facet_wrap(~name, dir="v")

```


### 4) Make an interactive plot of lakes in Iowa and Illinois and color them by lake area in hectares

```{r, echo=F}

#Plotting the first 1000 lakes
ii_lakes %>%
  slice(1:1000) %>%
  arrange(lake_area_ha) %>%
  mapview(.,zcol = 'lake_area_ha')

```


### 5) What other data sources might we use to understand how reservoirs and natural lakes vary in size in these three states? 

Other sources of data that we might use to explore the difference in lakes and reservoirs in the three starts are possible other public data forums. We could use base maps in arcgis. We could delineate watersheds on NWIS and use tools to calculate areas. 
